mlr_assg4

### 2022-09-21 hw4-1/abhinav/mlr.Rmd
Final Code push-2

```{r setup, include=FALSE}
library(cowplot)
library(ggplot2)
library(reshape2)
library(grid)
```

@@

* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place , error terms are normally distributed and residuals are homogeneous around the fitted line. However , we can see slight bends in the left and right end in the plot of residuals vs fitted values. Thus , a possible remedy can be thought of as modelling the log of dependent variable based on independent variable. (because taking a log transformation reduces the range of possible variables.)

# Points for further exploration  : 
* Taking a log transformation of the dependent variable and then create a regression model.
* Create a model with lower number of features (remove highly correlated features).
* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output.

@@

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```
Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here .\ 
Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers.

>>>>

```{r setup, include=FALSE}
library(cowplot)
library(ggplot2)
library(reshape2)
library(grid)
library(MASS)
```

@@

* From the 4 diagnostic plots associated with the linear regression model , we can see that the assumptions are mostly in place although there are some slight deviations, error terms are normally distributed (deviation for a couple of points) and residuals are mostly homogeneous around the fitted line. However , we can see slight bends in the left and right end in the plot of residuals vs fitted values. Thus , a possible remedy can be thought of as modelling the log of dependent variable based on independent variable. (because taking a log transformation reduces the range of possible variables.)

# Points for further exploration  : 
* Taking a log transformation of the dependent variable and then create a regression model.
* Create a model with lower number of features (remove highly correlated features).
* Look for possible outliers/leverage points (based on the residuals vs leverage plot - cook's distance) and remove them to see changes in the model output. 
* We can use Box-Cox transformation to identify the best possible transformation of y to regress on . This is usually done to handle the points deviating a bit from the normal Q-Q plot.

```{r}
# Box Cox Transformation (we can try this out to find the best possible function of y to regress on , the value of lambda is determined through MLE)
bc<-boxcox(lmodel_1, lambda =seq(-3,3))
#extract the best lambda 
best.lam<-bc$x[which(bc$y==max(bc$y))]

```
# Observations: 
Since , the above graph gives us a value of $\lambda \approx 0$ , therefore taking a log transformation of the dependent variable y makes more sense.

@@

new_datapoint<-data.frame(M=c(14.0),So=c(as.factor(0)),Ed=c(10.0),Po1=c(12.0),Po2=c(15.5),LF=c(0.640),M.F=c(94.0),Pop=c(150),NW=c(1.1),U1=c(0.120),U2=c(3.6),Wealth=c(3200),Ineq=c(20.1),Prob=c(0.04),Time=c(39.0))
crime_rate_pred<-predict(lmodel_2,newdata = new_datapoint)
crime_rate_pred<- exp(crime_rate_pred)
print(paste("Predicted Value:", round(crime_rate_pred,2)))
```
Although the $R^{2}$ takes a slight dip , but it appears that the model assumptions are better satisfied here .\ 
Points 11,19,46 seems to have around 1.5*standard deviation , thus these could be tested as potential outliers/influential points.
